---
title: "Trying to get ABC to run"
output: html_notebook
---

# Based on Thiele et al. 2014

This is a trial run of the ABC rejection sampling algorithm as described in *Thiele et al. 2014*. The code is adapted from SM9_abc_rejection.R of the supplementary material of that article.

## Getting rJava and RNetLogo to run

Connecting R, Java, and NetLogo through rJava and RNetLogo is not a trivial issue. Especially the compatibility of different R packages (R, Java, and NetLogo) is not always given: rJava may not recognize the current Java version, RNetLogo may not be compatible with an older NetLogo version, and rJava and RNetLogo may not be compatible.
Helpful links to connect the correct Java version may be found [here](http://stackoverflow.com/questions/26061726/compile-rjava-on-mac-for-java-1-7), [here](http://conjugateprior.org/2014/12/r-java8-osx/), [here](https://github.com/s-u/rJava/issues/37), [here](http://stackoverflow.com/questions/26618105/rnetlogo-not-working-on-mac-yosemite/26723881#26723881) or [here](http://stackoverflow.com/questions/13610293/how-to-get-rjava-0-9-3-to-work-on-os-x-10-7-4-with-oracle-java-1-7).
Older RNetLogo versions are available [here](https://cran.r-project.org/src/contrib/Archive/RNetLogo). For NetLogo version 5.0 - 5.2, [RNetLogo_1.0-1](https://cran.r-project.org/src/contrib/Archive/RNetLogo/RNetLogo_1.0-1.tar.gz) should be installed as follows. 
```{r, results='hide'}
packageurl <- "https://cran.r-project.org/src/contrib/Archive/RNetLogo/RNetLogo_1.0-1.tar.gz"
install.packages(packageurl, repos=NULL, type="source")
```


## Setup and run the simulation

This code block first loads all the necessary packages and draws the parameter sets following a latin hypercube sampling algorithm. Finally, the simulations are run based on the generated parameter set. 
```{r Setup and simulation}

#-------------------------------------------------------------------------------------
# I. Prerequirements
#-------------------------------------------------------------------------------------

# load RNetLogo package 
# (if not installed already, execute install.packages("rJava") and install.packages("RNetLogo") )
Sys.setenv(NOAWT=1)
require(RNetLogo)

# an R random seed (for beeing reproducible)
set.seed(1606350664)

# TODO: adapt these paths to your requirements
# the NetLogo installation path (where the NetLogo.jar is located)
nl.path     <- "/Applications/NetLogo 5.2"
# the path to the NetLogo model file
# model.path  <- "../../src/irrigame_comm.nlogo"
model.path  <- "/Users/felixx/Documents/UFZ-Cloud/IrriGameABM/src/irrigame_comm.nlogo"
# the simulation function
# simfun.path <- "../../src/sim_function1.R"
simfun.path <- "/Users/felixx/Documents/UFZ-Cloud/IrriGameABM/src/sim_function1.R"

# TODO: set values for those parameters that are not changes during the experiment
#       parameter names have to be the names of the parameters in the NetLogo model
fix.params <- list(
                    'scenario'      = "\"social-values\"",
                    'limcom'        = "false",
                    'visioneffect'  = "false",
                    'phase2?'       = "false"
)      

# TODO: give value ranges for each parameter 
#       parameter names has to been the names of the parameters in the NetLogo model
parameter.values <- list(
                          'psat'          = list(min = 0.0,  max = 1.0, random.function="qunif"),
                          'umin'          = list(min = 0.0,  max = 50.0, random.function="qunif"),
                          'alpha'         = list(min = -1.0, max = 1.0, random.function="qunif"),
                          'beta'          = list(min = -1.0, max = 1.0, random.function="qunif"),
                          'mu'            = list(min = 0.0,  max = 100.0, random.function="qunif")
                        )

# TODO: adapt to your observation summary statistics (here: just the mean of the ranges)
calibration.ranges <- data.frame(
                            'Infrastructure2' = 83.38,
                            'Infrastructure4' = 78.43, 
                            'Infrastructure6' = 74.19,
                            'Infrastructure8' = 74.00,
                            'Infrastructure10' = 72.57,
                            'Inequality2' = 0.283,
                            'Inequality4' = 0.223, 
                            'Inequality6' = 0.241, 
                            'Inequality8' = 0.182, 
                            'Inequality10' = 0.257
                          )

# TODO: number of parameter sets
sample.count <- 30000

# how many repetitions for each parameter set should be run (to control stochasticity)?
# TODO: adapt the number of repititions, set to 1 if deterministic model
no.repeated.sim <- 10

# TODO: set tolerance value for ABC (required proportion of points accepted)
tol <- 0.3

# TODO: should R report the progress
trace.progress = TRUE

# initialize NetLogo
nl.sm9 <- "irrigame_comm.nlogo"
NLStart(nl.path, gui=FALSE, nl.obj=nl.sm9)
NLLoadModel(model.path,nl.obj=nl.sm9)
# NLQuit(all = T)

#-------------------------------------------------------------------------------------
# II. Definition of the simulation function to test a parameter set
#-------------------------------------------------------------------------------------

# load the code of the simulation function (name: simulate)
source(file=simfun.path)


#-------------------------------------------------------------------------------------
# III. Creation of the parameter sets by latin hypercube sampling
#-------------------------------------------------------------------------------------

# get names of parameters
parameter.names <- names(parameter.values)
fix.param.names <- names(fix.params)

#-------------------------------------------------------------------------------------
# III.A. Alternative 1: Package tgp
#-------------------------------------------------------------------------------------
# load required package tgp
#require(tgp)

# create random (latin hypercube) samples, it would be possible to use the ranges directly
# by we will use the value interval [0,1] and apply a random distribution function afterwards
#lhs.design <- lhs(sample.count, t(matrix(c(0,1),2,length(parameter.values))))

# transform the standardized random values to the real parameter value range
# and apply the desired random distribution
#lhs.design <- lapply(seq(1,length(parameter.values)), function(i) {
#    match.fun(parameter.values[[i]]$random.function)(lhs.design[,i], parameter.values[[i]]$min, parameter.values[[i]]$max)
#  })
#names(lhs.design) <- parameter.names


#-------------------------------------------------------------------------------------
# III.B. Alternative 2: Package lhs
#-------------------------------------------------------------------------------------
# NOTE: here, we can't give the real value range. 
# This package will create values from 0 to 1 and must be transformed afterwards.
# But this package is much more powerfull supplying different LHS methods.
require(lhs)

# create a random sample of parameter sets (Latin Hypercube Sampling)
lhs.design <- randomLHS(sample.count, length(parameter.values))

# transform the standardized random values to the real parameter value range
# and apply the desired random distribution
lhs.design <- lapply(seq(1,length(parameter.values)), function(i) {
    match.fun(parameter.values[[i]]$random.function)(lhs.design[,i], parameter.values[[i]]$min, parameter.values[[i]]$max)
  })
names(lhs.design) <- parameter.names


#-------------------------------------------------------------------------------------
# IV. Run of the (LHS) simulation for all parameter sets
#-------------------------------------------------------------------------------------

# variable used for progress tracing 
# (global variable: not nice, but I see currently no other way in R)
already.processed <- 0

# simulate for all parameter sets and
# get results of all evalulation criteria as matrix
sim.results.lhs <- apply(data.frame(lhs.design), 1, 
                         simulate,
                         no.repeated.sim=no.repeated.sim, nl.obj=nl.sm9, 
                         trace.progress=trace.progress,
                         parameter.names=parameter.names, 
                         iter.length=sample.count,
                         function.name="LHS",
                         fix.params, fix.param.names
                        )
rownames(sim.results.lhs) <- c(sapply(c(2,4,6,8,10), function(x) {paste0("Infrastructure", x)}),
                            sapply(c(2,4,6,8,10), function(x) {paste0("gini_ext", x)}))

```

## Data analysis

### Baseline scenario

The following code block runs the ABC analysis.
```{r ABC setup}
#-------------------------------------------------------------------------------------
# V. Run the ABC
#-------------------------------------------------------------------------------------

# load required package abc
require(abc)

# reorganize the data
sumstat <- as.data.frame(t(sim.results.lhs))
names(sumstat) <- names(calibration.ranges)
param <- as.data.frame(lhs.design)

# covariance calculation to get a feeling for a good tolerance value
cv.rej <- cv4abc(param=param, sumstat=sumstat, nval=50,
                 tols=c(.05,.1,.2,.3,.5), method="rejection")
summary(cv.rej)
plot(cv.rej, log=rep("xy", times = length(param)))

```

```{r}
# run the ABC with rejection sampling
method <- "rejection"
sim.results.abc.rej <- abc(target=calibration.ranges, param=param, sumstat=sumstat, tol=tol, method=method)

# plot the resulting posterior distribution for the parameters
# (package abc must be loaded)
hist(sim.results.abc.rej, breaks=50)

```

```{r}
# run the ABC with local linear regression method
method <- "loclinear" # "neuralnet" (require(nnet))
sim.results.abc.ll <- abc(target=calibration.ranges, param=param, sumstat=sumstat, tol=tol, method=method)

# create diagnostic plot
# (package abc must be loaded)
plot(sim.results.abc.ll, param)

# plot the resulting posterior distribution for the parameters
# (package abc must be loaded)
hist(sim.results.abc.ll, breaks=50)
```

```{r}
# distribution characteristics
summary(sim.results.abc.rej,intvl=0.9)
summary(sim.results.abc.ll,intvl=0.9)
```

### Keeping only the 10% best performing runs

```{r}
# set the tolerance threshold to 10%
tol <- 0.1 


# run the ABC with rejection sampling
method <- "rejection"
sim.results.abc.rej <- abc(target=calibration.ranges, param=param, sumstat=sumstat, tol=tol, method=method)

# plot the resulting posterior distribution for the parameters
# (package abc must be loaded)
hist(sim.results.abc.rej, breaks=50)
```


```{r}
# run the ABC with local linear regression method
method <- "loclinear" # "neuralnet" (require(nnet))
sim.results.abc.ll <- abc(target=calibration.ranges, param=param, sumstat=sumstat, tol=tol, method=method)

# create diagnostic plot
# (package abc must be loaded)
plot(sim.results.abc.ll, param)

# plot the resulting posterior distribution for the parameters
# (package abc must be loaded)
hist(sim.results.abc.ll, breaks=50)
```

```{r}
# distribution characteristics
summary(sim.results.abc.rej,intvl=0.9)
summary(sim.results.abc.ll,intvl=0.9)
```


### Calibrating against all time steps

```{r Setup and simulation, results='hide'}

#-------------------------------------------------------------------------------------
# I. Prerequirements
#-------------------------------------------------------------------------------------

# load RNetLogo package 
# (if not installed already, execute install.packages("rJava") and install.packages("RNetLogo") )
Sys.setenv(NOAWT=1)
require(RNetLogo)

# an R random seed (for beeing reproducible)
set.seed(1606350664)

# TODO: adapt these paths to your requirements
# the NetLogo installation path (where the NetLogo.jar is located)
nl.path     <- "/Applications/NetLogo 5.2"
# the path to the NetLogo model file
# model.path  <- "../../src/irrigame_comm.nlogo"
model.path  <- "/Users/felixx/Documents/UFZ-Cloud/IrriGameABM/src/irrigame_comm.nlogo"
# the simulation function
# simfun.path <- "../../src/sim_function1.R"
simfun.path <- "/Users/felixx/Documents/UFZ-Cloud/IrriGameABM/src/sim_function1.R"

# TODO: set values for those parameters that are not changes during the experiment
#       parameter names have to be the names of the parameters in the NetLogo model
fix.params <- list(
                    'scenario'      = "\"social-values\"",
                    'limcom'        = "false",
                    'visioneffect'  = "false",
                    'phase2?'       = "false"
)      

# TODO: give value ranges for each parameter 
#       parameter names has to been the names of the parameters in the NetLogo model
parameter.values <- list(
                          'psat'          = list(min = 0.0,  max = 1.0, random.function="qunif"),
                          'umin'          = list(min = 0.0,  max = 50.0, random.function="qunif"),
                          'alpha'         = list(min = -1.0, max = 1.0, random.function="qunif"),
                          'beta'          = list(min = -1.0, max = 1.0, random.function="qunif"),
                          'mu'            = list(min = 0.0,  max = 100.0, random.function="qunif")
                        )

# TODO: adapt to your observation summary statistics (here: just the mean of the ranges)
calibration.ranges <- data.frame(
                            'Infrastructure1' = 92.95,
                            'Infrastructure2' = 83.38,
                            'Infrastructure3' = 79.14, 
                            'Infrastructure4' = 78.43,
                            'Infrastructure5' = 78.43,
                            'Infrastructure6' = 74.19,
                            'Infrastructure7' = 73.76,
                            'Infrastructure8' = 74.00,
                            'Infrastructure9' = 74.24,
                            'Infrastructure10' = 72.57,
                            'Inequality1' = 0.291,
                            'Inequality2' = 0.283,
                            'Inequality3' = 0.280,
                            'Inequality4' = 0.223, 
                            'Inequality5' = 0.249,
                            'Inequality6' = 0.241, 
                            'Inequality7' = 0.264,
                            'Inequality8' = 0.182, 
                            'Inequality9' = 0.219,
                            'Inequality10' = 0.257
                          )

# TODO: number of parameter sets
sample.count <- 30000

# how many repetitions for each parameter set should be run (to control stochasticity)?
# TODO: adapt the number of repititions, set to 1 if deterministic model
no.repeated.sim <- 10

# TODO: set tolerance value for ABC (required proportion of points accepted)
tol <- 0.3

# TODO: should R report the progress
trace.progress = TRUE

# initialize NetLogo
nl.sm9 <- "irrigame_comm.nlogo"
NLStart(nl.path, gui=FALSE, nl.obj=nl.sm9)
NLLoadModel(model.path,nl.obj=nl.sm9)
# NLQuit(all = T)

#-------------------------------------------------------------------------------------
# II. Definition of the simulation function to test a parameter set
#-------------------------------------------------------------------------------------

# load the code of the simulation function (name: simulate)
source(file=simfun.path)


#-------------------------------------------------------------------------------------
# III. Creation of the parameter sets by latin hypercube sampling
#-------------------------------------------------------------------------------------

# get names of parameters
parameter.names <- names(parameter.values)
fix.param.names <- names(fix.params)

#-------------------------------------------------------------------------------------
# III.A. Alternative 1: Package tgp
#-------------------------------------------------------------------------------------
# load required package tgp
#require(tgp)

# create random (latin hypercube) samples, it would be possible to use the ranges directly
# by we will use the value interval [0,1] and apply a random distribution function afterwards
#lhs.design <- lhs(sample.count, t(matrix(c(0,1),2,length(parameter.values))))

# transform the standardized random values to the real parameter value range
# and apply the desired random distribution
#lhs.design <- lapply(seq(1,length(parameter.values)), function(i) {
#    match.fun(parameter.values[[i]]$random.function)(lhs.design[,i], parameter.values[[i]]$min, parameter.values[[i]]$max)
#  })
#names(lhs.design) <- parameter.names


#-------------------------------------------------------------------------------------
# III.B. Alternative 2: Package lhs
#-------------------------------------------------------------------------------------
# NOTE: here, we can't give the real value range. 
# This package will create values from 0 to 1 and must be transformed afterwards.
# But this package is much more powerfull supplying different LHS methods.
require(lhs)

# create a random sample of parameter sets (Latin Hypercube Sampling)
lhs.design <- randomLHS(sample.count, length(parameter.values))

# transform the standardized random values to the real parameter value range
# and apply the desired random distribution
lhs.design <- lapply(seq(1,length(parameter.values)), function(i) {
    match.fun(parameter.values[[i]]$random.function)(lhs.design[,i], parameter.values[[i]]$min, parameter.values[[i]]$max)
  })
names(lhs.design) <- parameter.names


#-------------------------------------------------------------------------------------
# IV. Run of the (LHS) simulation for all parameter sets
#-------------------------------------------------------------------------------------

# variable used for progress tracing 
# (global variable: not nice, but I see currently no other way in R)
already.processed <- 0

# simulate for all parameter sets and
# get results of all evalulation criteria as matrix
sim.results.lhs <- apply(data.frame(lhs.design), 1, 
                         simulate_all,
                         no.repeated.sim=no.repeated.sim, nl.obj=nl.sm9, 
                         trace.progress=trace.progress,
                         parameter.names=parameter.names, 
                         iter.length=sample.count,
                         function.name="LHS",
                         fix.params, fix.param.names
                        )
rownames(sim.results.lhs) <- c(sapply(1:10, function(x) {paste0("Infrastructure", x)}),
                            sapply(1:10, function(x) {paste0("gini_ext", x)}))

```



The following code block runs the ABC analysis.
```{r ABC setup}
#-------------------------------------------------------------------------------------
# V. Run the ABC
#-------------------------------------------------------------------------------------

# load required package abc
require(abc)

# reorganize the data
sumstat <- as.data.frame(t(sim.results.lhs))
names(sumstat) <- names(calibration.ranges)
param <- as.data.frame(lhs.design)

# covariance calculation to get a feeling for a good tolerance value
cv.rej <- cv4abc(param=param, sumstat=sumstat, nval=50,
                 tols=c(.05,.1,.2,.3,.5), method="rejection")
summary(cv.rej)
plot(cv.rej, log=rep("xy", times = length(param)))

```

```{r}
# run the ABC with rejection sampling
method <- "rejection"
sim.results.abc.rej <- abc(target=calibration.ranges, param=param, sumstat=sumstat, tol=tol, method=method)

# plot the resulting posterior distribution for the parameters
# (package abc must be loaded)
hist(sim.results.abc.rej, breaks=50)

```

```{r}
# run the ABC with local linear regression method
method <- "loclinear" # "neuralnet" (require(nnet))
sim.results.abc.ll <- abc(target=calibration.ranges, param=param, sumstat=sumstat, tol=tol, method=method)

# create diagnostic plot
# (package abc must be loaded)
plot(sim.results.abc.ll, param)

# plot the resulting posterior distribution for the parameters
# (package abc must be loaded)
hist(sim.results.abc.ll, breaks=50)
```

```{r}
# distribution characteristics
summary(sim.results.abc.rej,intvl=0.9)
summary(sim.results.abc.ll,intvl=0.9)
```

### Keeping only the 10% best performing runs

```{r}
# set the tolerance threshold to 10%
tol <- 0.1 


# run the ABC with rejection sampling
method <- "rejection"
sim.results.abc.rej <- abc(target=calibration.ranges, param=param, sumstat=sumstat, tol=tol, method=method)

# plot the resulting posterior distribution for the parameters
# (package abc must be loaded)
hist(sim.results.abc.rej, breaks=50)
```


```{r}
# run the ABC with local linear regression method
method <- "loclinear" # "neuralnet" (require(nnet))
sim.results.abc.ll <- abc(target=calibration.ranges, param=param, sumstat=sumstat, tol=tol, method=method)

# create diagnostic plot
# (package abc must be loaded)
plot(sim.results.abc.ll, param)

# plot the resulting posterior distribution for the parameters
# (package abc must be loaded)
hist(sim.results.abc.ll, breaks=50)
```

```{r}
# distribution characteristics
summary(sim.results.abc.rej,intvl=0.9)
summary(sim.results.abc.ll,intvl=0.9)
```